### 2021-10-13 plos_latex_template.tex
Rocío: shortening rule 1

Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. 
However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.
An exclusionary conference experience can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018}.
Efforts to address systematic discrimination in conferences and make them inclusive spaces can have a meaningful impact on our communities and the lives and careers of their members.

This article suggests rules to pivot traditional conferences towards diversity and inclusion, and strive to build more inclusive and welcoming communities. 
Some scholarly or technological communities are formally structured as learned societies (e.g. the International Geographical Union or the Royal Statistical Society), others are composed by networks of less formal local groups (e.g. R User Groups or Python User Groups), and others may not have any established structure. 
For the latter, conferences may play even a bigger role in shaping the community, since there is no other organization or institution to gather the community members.

The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
The rules can also be helpful to local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
These tips stem from the authors' experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for everyone. 
Here, we share the lessons learned within the past year of organizing this global useR! 2021, summarized as ten simple rules towards an inclusive conference.
We have organized these rules in three groups.
Rules 1, 2 and 10 refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
The next two rules are focused on the people that participate in the conference: 
Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of removing implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs. 
Rules 5 to 9 are rules about actions regarding components of the conference that should be carefully planned for: an online component, accessibility to people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion. 
These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.
% Rocío: the last sentence was copied from the abstract because we need a closing sentence.
% Rocío: Could add a line introducing Fig. in https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing and the whole structure [x]

@@

Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few. 
Human diversity should be celebrated and respected in every way. Nonetheless, we live in a world with implicit hierarchies along these axes. Some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems--including conferences--are consciously and unconsciously built.  
While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.

To be proactive allies in our conferences and communities, we should start by examining our own privileges (unearned advantages given by society to some people but not all).
Privilege is largely invisible to those who have it. And we will not be able to help end discrimination and oppression if we do not recognize our privileges first and the injustices that come with them.
Once we understand our own privileges, we can begin addressing them on an individual, collective, and institutional basis.
Recognizing our privileges--particularly in our field and in our scientific or professional community--will help identify which subgroups have been the most excluded or discriminated against. 
These are the groups we need to make more effort to include. 
Depending on the field, region, or community, it could be Black People, LGBTQIA+, Muslim, or others. 
%ast: ö check phrasing, add "people"
Intersectionality, or the complex, cumulative way in which the effects of multiple forms of discrimination combine or intersect, should be taken into account.

Investing more effort in the most excluded groups does not mean neglecting the others (see \textbf{Rule \ref{rule_inclusion}} for more on that). 
But it will prevent from falling into the mistake of treating diversity as a checklist.
It will also guide the vision of diversity for your conference--and your strategies to achieve it:
Will a more diverse conference translate into an even gender distribution in your speakers? 
Would it be the presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees? 
Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
This exercise will also help coming up with indicators that would help the organizing team assess the fulfillment of the diversity objectives along the way.

>>>>

% Rocío: Main message of the paragraph: Why we are writing this piece
Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. 
However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.
An exclusionary conference experience can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018}.
Efforts to address systematic discrimination in conferences and make them inclusive spaces can have a meaningful impact on our communities and the lives and careers of their members.

% Rocío: Main message of the paragraph: We do this for communities and people. What do we mean by communities and why are they important? (Why are they important may need some text)
This article suggests rules to pivot traditional conferences towards diversity and inclusion, and strive to build more inclusive and welcoming communities. 
Some scholarly or technological communities are formally structured as learned societies (e.g. the International Geographical Union or the Royal Statistical Society), others are composed by networks of less formal local groups (e.g. R User Groups or Python User Groups), and others may not have any established structure. 
For the latter, conferences may play even a bigger role in shaping the community, since there is no other organization or institution to gather the community members.

% Rocío: Main message of the paragraph: Roadmap of the paper
The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
The rules can also be helpful to local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
These tips stem from the authors' experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for everyone. 
Here, we share the lessons learned within the past year of organizing this global useR! 2021, summarized as ten simple rules towards an inclusive conference.
We have organized these rules in three groups.
Rules 1, 2 and 10 refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
The next two rules are focused on the people that participate in the conference: 
Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of removing implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs. 
Rules 5 to 9 are rules about actions regarding components of the conference that should be carefully planned for: an online component, accessibility to people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion. 
These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.
% Rocío: the last sentence was copied from the abstract because we need a closing sentence.
% Rocío: Could add a line introducing Fig. in https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing and the whole structure [x]

@@

% Rocío: Main idea: what is diversity and why does it matter? (inequalities)
Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few. 
Human diversity should be celebrated and respected in every way. Nonetheless, we live in a world with implicit hierarchies along these axes. Some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems--including conferences--are consciously and unconsciously built.  
While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.

% Rocío: more effort towards including the more excluded
Recognizing our privileges--unearned advantages given by society to some people but not all--particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against. 
These are the groups we need to make more effort to include. 
Investing more effort in the most excluded groups does not mean neglecting the others (see \textbf{Rule \ref{rule_inclusion}} for more on that), but it will guide the vision of diversity for your conference--and your strategies to achieve it:
Will a more diverse conference translate into an even gender distribution in your speakers? 
Would it be the presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees? 
Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
Depending on the field, region, or community, it could be Black people, LGBTQIA+, Muslim, or other people.
Intersectionality, or the complex, cumulative way in which the effects of multiple forms of discrimination combine or intersect, should be taken into account. 
% This exercise will also help coming up with indicators that would help the organizing team assess the fulfillment of the diversity objectives along the way.
